---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# TGIFA

<!-- badges: start -->
<!-- badges: end -->

The TGIFA R package presents a set of functions that perform missing data imputation using a truncated Gaussian infinite factor analysis model designed for high-dimensional metabolomics data.

## Installation

You can install the development version of TGIFA like so:

``` r
# install remotes package if not already installed
# install.packages("remotes")
remotes::install_github("kfinucane/TGIFA")

# load the library
library(TGIFA)

```

## Example imputation

To illustrate the TGIFA functionality, here a test dataset is generated with two missing values.

```{r, results = FALSE}
# load the library
library(TGIFA)

# generate example data
set.seed(1)
example_data <- matrix(abs(rnorm(200)), nrow = 5)

# add missingness to example data coded as 0.001
example_data[4, 2] <- 0.001
example_data[2, 18] <- 0.001
```

This example data can then be passed into the `TGIFA_model()` function which contains all package functionality. Here, `input_data` is the dataset with missing values, in matrix format. The `coding` argument refers to how the missing values are coded. If your missing values all have a value of `NA`, for example, you should input `coding = NA`. Here, for the example, missing entries are coded with the value `0.001`. Here, the value `k.star = 3` refers to the practical non-infinite number of latent factors used by the TGIFA model; usually this defaults to `k.star = 5` but for this small example dataset we will use a smaller number. 

The remaining parameters refer to the MCMC procedure; here `n.iters = 100` runs the MCMC chain for 100 iterations, with `burn = 50` controlling the number of draws discarded in a burn and `thin = 5` stating that every fifth post-burn draw from the MCMC chain should be retained. Please note that `burn` should have a value less than `n.iters`, and `thin` should divide into both `burn`, and `n.iters` - `burn` with no remainder.

```{r, results = FALSE}
# run TGIFA model
# short chain for example
res <- TGIFA_model(input_data = example_data, coding = 0.001, n.iters = 100, k.star = 3,
                  burn = 50, thin = 5)
```

The results of the TGIFA imputation can be accessed as follows.


```{r, results = FALSE}
# checkout imputed dataset
res$imputed_dataset

# checkout further information on imputed entries
res$imputation_info
```

The input dataset, with missing values imputed according to the TGIFA imputation method is contained in `res$imputed_dataset` in matrix format. Details on the imputation can be accessed from `res$imputation_info`. This is a dataframe with number of rows equal to the number of missing values in the dataset, and seven columns. The `entry_row` and `entry_col` columns contain the row and column index of each missing value in the input dataset. The `imputed_val` column contains the imputed value, and `cred_int_upper` and `cred_int_lower` contain the upper and lower limits of hte 95\% credible interval for the imputed point. Finally, the `miss_mech` column gives the final inferred missingness type of each point, with `miss_mech_unc` providing the corresponding uncertainty.

If one wishes to change the hyperparameters of the TGIFA model from their defaults, this can be done as follows. All parameters are as given in the TGIFA paper.

```{r, eval = FALSE}
res <- TGIFA_model(input_data, coding = NA, n.iters = 10000, k.star = 5,
                   verbose = TRUE, burn = 5000, thin = 5,
                   kappa_1 = 3L, kappa_2 = 2L,
                   a_sigma = 1L, b_sigma = 0.25, a_1 = 2.1, a_2 = 3.1,
                   prop_mult = list("mu_prop_mult" = 0.05,
                                    "eta_prop_mult" = 3L,
                                    "lambda_prop_mult" = 3L,
                                    "sigma_inv_prop_mult" = 1.5),
                   return_chain = FALSE,
                   cred_level = 0.95,
                   var_mult = 1.0)
```
